Future challenges for Research Software Engineering

### 2025-05-06 future_challenges_for_RSEng.tex
small rephrase formulated mostly by Julian
Co-authored-by: Julian Dehne <julian.dehne@rz.gi.de>

\subsection{Ethics and social consideration}
The ethics and values of the Research Software Engineer are grounded in their practice.
They commit to the health, safety and welfare of the public and act in the interest of society, their employer and their clients \cite{Goth2023}.
Sometimes these values are in direct conflict with the aims of their institution.
For example, their PI might insist on quickly developing some software without adhering to current best practices.
Another example might be that they are developing code that can also be used for military purposes, such as numerical solvers or computer vision systems.
Like engineers, RSEs take the responsibility for the software they develop.
RSEs need to be aware of these issues and develop their own ``red lines''.

RSEs are publicly funded.
Similarly to how data collected by scientists should adhere to the FAIR principles \cite{FAIR}, research software developed by RSEs should adhere to the FAIR4RS principles~\cite{FAIR4RS}.

>>>>

\subsection{Ethics and social consideration}

The ethics and values of the Research Software Engineer are grounded in their practice.
They commit to the health, safety and welfare of the public and act in the interest of society, their employer and their clients \cite{Goth2023}.

% institutional bonds and ethics
Sometimes these values are in direct conflict with the aims of their institution. For example, their PI might insist on quickly developing some software without adhering to current best practices.
A code of ethics and professional conduct for RSEs should be developed by institutions together with RSEs themselves. This can then form the basis for institutional policies and guidelines.

% society implications
The technology people working as RSE might have, well directly or indirectly, ethical implications in areas such as social engineering [TODO find source], climate [TODO find sourc] and more.
Another example might be that they are developing code that can also be used for military purposes, such as numerical solvers or computer vision systems.
Like engineers, RSEs take the responsibility for the software they develop.
RSEs need to be aware of these issues and develop their own ``red lines''.

% data ethics
Similarly to how data collected by scientists should adhere to the FAIR principles \cite{FAIR}, research software developed by RSEs should adhere to the FAIR4RS principles~\cite{FAIR4RS}.
